---
title: "npnSpline"
author: "K. Duffy"
date: "October 18, 2018"
output: html_document
---

Start my loading some npn data, we had to do a backdoor of this so I won't include code here

Filter the data to just grab 'leaves' data

```{r}
leaves_idx=is.element(status_intensity_observation_data$Phenophase_Description, 'Leaves')
leaves=status_intensity_observation_data[leaves_idx,]
leaves=leaves%>%filter(Phenophase_Status>0)
#idx=leaves$Phenophase_Status>0
```

Aggregate to count data:
```{r}
count_data=aggregate.data.frame(x=leaves$Phenophase_Status, by=list(leaves$Observation_Date), FUN=sum)
```

Plot the data to undertand what we are working with:
```{r}
p=ggplot(count_data, aes(x=Group.1, y=x))+
  geom_point(alpha=.5, color='darkgreen') #this sets the transparency value
p +theme_bw() #this just cleans up the base grey background from ggplot2
```

```{r}
library(stats)
library(ggthemes)
library(ggplot2)
s=smooth.spline(x=count_data) #here is an ultra basic smoothing spline, there are a bunch of additional parameters you can include to trailor the smoothing.  probably easiest to just chat about it
p=ggplot(count_data, aes(x=Group.1, y=x))+
  geom_point(alpha=.2, color='darkgreen')+
  geom_line(aes(x=Group.1, y=s$y), color='darkgreen')
p+ggthemes::theme_few()

```

A LOESS spline, where you can set a proportion of that data to smooth based on (span below)
```{r}
count_data$Group.1=as.numeric(count_data$Group.1) #loess smoothes don't accept date format
l=loess(x ~ Group.1, data=count_data, span=.3) #, span=0.10

p=ggplot(count_data, aes(x=Group.1, y=x))+
  geom_point(alpha=.2, color='darkgreen')+
  geom_line(aes(x=Group.1, y=l$fitted), color='darkgreen')
p+theme_bw()
```
Grab the error associated with the models:

```{r}
library(Metrics)
error_spline=rmse(count_data$x, s$y )
error_spline  
error_LOESS=rmse(count_data$x, l$fitted )
error_LOESS
```
Basically the error is the same, so really you could pick either!  We can talk more about this soon but I'll end here because I have to jump on a plane :)